SSwAB-IT Phase 1 Analysis

### 2024-03-21 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
SSwAB-IT phase 1 excluded positive/positive

---
title: "SSwAB-IT Phase 1 Analysis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly

editor: source
---

```{r setup}
#| include: false
library(redcapAPI)
library(tidyverse)
library(gt)
library(gtsummary)
library(caret)
```

@@

```{r}
df %>%
  select(all_of(bugs)) %>%
  tbl_summary(label = bug_list, missing_text = "Missing") %>%
  add_ci(statistic=list(all_categorical() ~ "{conf.low} - {conf.high}%"))
```

```{r}
agreement <- function(bug, data = df) {
  cos <- paste0(bug, "_cos")
  sos <- paste0(bug, "_sos")
  d <- data[, c(cos, sos)]
  d <- d[complete.cases(d), ]
  # Agreement metrics
  c <- confusionMatrix((d[, sos]),
    reference = d[, cos], positive = "Positive", mode = "everything"
  )
  # Return
  return(list(two_by_two = c$table, accuracy = c$overall, metrics = c$byClass))
}
```

```{r}
l <- lapply(bugs, function(b) {
  a <- agreement(b)
  return(list(
    Accuracy = a$accuracy["Accuracy"],
    `Accuracy Lower Limit` = a$accuracy["AccuracyLower"],
    `Accuracy Upper Limit` = a$accuracy["AccuracyUpper"],
    Kappa = a$accuracy["Kappa"], `Mcnemar P Value` = a$accuracy["McnemarPValue"]
  ))
})
l <- data.frame(do.call(rbind, l), check.names = F)
l <- data.frame(lapply(l, as.numeric), check.names = F)
rownames(l) <- bug_names
gt(l, rownames_to_stub = T) %>% fmt_number(n_sigfig = 3)
```

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("##", as.character(bug_list[b]))
  cat("\n")
  cos <- paste0(b, "_cos")
  sos <- paste0(b, "_sos")
  d <- df[, c(cos, sos)]
  d <- d[complete.cases(d), ]
  cat("\n")
  c <- confusionMatrix(d[, sos],
    reference = d[, cos], positive = "Positive", mode = "everything"
  )
  metrics <- data.frame(c$overall)
  metrics2 <- data.frame(c$byClass)
  cat("### 2x2 table")
  cat("\n")
  two_by_two <- data.frame(c$table)
  colnames(two_by_two) <- c("SOS", "COS", "n")
  print(gt(two_by_two))
  cat("\n")
  cat("### Metrics")
  cat("\n")
  print(gt(metrics, rownames_to_stub = T) %>%
    fmt_number(n_sigfig = 3) %>%
    tab_options(column_labels.hidden = TRUE))
  cat("\n")
  print(gt(metrics2, rownames_to_stub = T) %>%
    fmt_number(n_sigfig = 3) %>%
    tab_options(column_labels.hidden = TRUE))
  cat("\n")
}))
```

# Questions

1. Are we okay with calculating prevalence from the data for the PPV and NPV? Do these numbers seem pretty representative of what you'd expect in the population?

>>>>

---
title: "SSwAB-IT Phase 1 Analysis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Documents/Notes/zotero.bib
csl: /Users/timvigers/GitHub/styles/american-medical-association-10th-edition.csl
editor: source
---

```{r setup}
#| include: false
library(redcapAPI)
library(tidyverse)
library(gt)
library(gtsummary)
library(caret)
library(vcd)
library(eulerr)
```

@@

For these agreement analyses, we reported accuracy metrics that include all samples. Accuracy is the proportion of samples where the two tests agreed (positive/positive and negative/negative). Cohen's kappa accounts for agreement between the two tests by chance, and is interpreted analagously to correlation where -1 indicates perfect disagreement and 1 indicates perfect agreement. Cohen's kapps of NaN or NA indicates a lack of variability between the two tests.

| Cohen's Kappa | Interpretation         |
|---------------|------------------------|
| 0             | No agreement           |
| 0.10 - 0.20   | Slight agreement       |
| 0.21 - 0.40   | Fair agreement         |
| 0.41 - 0.60   | Moderate agreement     |
| 0.61 - 0.80   | Substantial agreement  |
| 0.81 - 0.99   | Near perfect agreement |
| 1             | Perfect agreement      |

Bangdiwala's agreement plot @bangdiwalaAgreementChart2013 is a way of visualizing 2 x 2 contingency tables, and additional details are in the paper referenced at the end of this document.

```{r}
agreement <- function(bug, data = df) {
  cos <- paste0(bug, "_cos")
  sos <- paste0(bug, "_sos")
  d <- data[, c(cos, sos)]
  d <- d[complete.cases(d), ]
  # Agreement metrics
  c <- confusionMatrix((d[, sos]),
    reference = d[, cos], positive = "Positive", mode = "everything",
    dnn = c("SOS", "COS")
  )
  # Return
  return(list(two_by_two = c$table, accuracy = c$overall))
}
```

```{r}
l <- lapply(bugs, function(b) {
  a <- agreement(b)
  # Remove those where both tests were positive
  n <- a$two_by_two
  n[2, 2] <- 0
  # Proportion tests for non-agreement in participants with one negative test
  cos_p <- prop.test(n[1, 2], sum(n))
  sos_p <- prop.test(n[2, 1], sum(n))
  nn <- prop.test(n[1, 1], sum(n))
  return(list(
    Accuracy = a$accuracy["Accuracy"],
    `Accuracy LL` = a$accuracy["AccuracyLower"],
    `Accuracy UL` = a$accuracy["AccuracyUpper"],
    Kappa = a$accuracy["Kappa"],
    `Mcnemar P Value` = a$accuracy["McnemarPValue"],
    `N With Negative Test on Either` = sum(n),
    `N Negative on Both` = n[1, 1],
    `Prop. Both Negative` = nn$estimate,
    `Prop. Both Negative LL ` = nn$conf.int[1],
    `Prop. Both Negative UL ` = nn$conf.int[2],
    `N Postive on COS and Negative on SOS` = n[1, 2],
    `Prop. Postive on COS` = cos_p$estimate,
    `Prop. Postive on COS LL ` = cos_p$conf.int[1],
    `Prop. Postive on COS UL ` = cos_p$conf.int[2],
    `N Postive on SOS and Negative on COS` = n[2, 1],
    `Prop. Postive on SOS` = sos_p$estimate,
    `Prop. Postive on SOS LL ` = sos_p$conf.int[1],
    `Prop. Postive on SOS UL ` = sos_p$conf.int[2]
  ))
})
l <- data.frame(do.call(rbind, l), check.names = F)
l <- data.frame(lapply(l, as.numeric), check.names = F)
rownames(l) <- bug_names
gt(l, rownames_to_stub = T) %>% fmt_number(n_sigfig = 3)
```

```{r results='asis'}
invisible(lapply(names(bug_list), function(b) {
  cat("\n")
  cat("##", as.character(bug_list[b]))
  cat("\n")
  cos <- paste0(b, "_cos")
  sos <- paste0(b, "_sos")
  d <- df[, c(cos, sos)]
  d <- d[complete.cases(d), ]
  colnames(d) <- c("COS", "SOS")
  cat("\n")
  c <- confusionMatrix(d[, "SOS"],
    reference = d[, "COS"], positive = "Positive", mode = "everything"
  )
  metrics <- data.frame(c$overall)
  cat("### 2x2 table")
  cat("\n")
  two_by_two <- data.frame(c$table)
  print(gt(two_by_two))
  cat("\n")
  cat("### Accuracy")
  cat("\n")
  print(gt(metrics, rownames_to_stub = T) %>%
    fmt_number(n_sigfig = 3) %>%
    tab_options(column_labels.hidden = TRUE))
  cat("\n")
  cat("### Agreement plot")
  cat("\n")
  print(agreementplot(table(d)))
  cat("\n")
  cat("### Positive/Positive Excluded")
  cat("\n")
  t <- d[rowSums(d == "Positive") < 2, ]
  t$`Samples With a Negative Test` <- paste(
    "COS", t[, "COS"], "/", "SOS", t[, "SOS"]
  )
  t %>%
    select(`Samples With a Negative Test`) %>%
    tbl_summary() %>%
    add_ci() %>%
    as_gt() %>%
    print()
  cat("\n")
  cat("### Euler diagram of negative results")
  cat("\n")
  print(plot(euler(d == "Negative"), quantities = T))
  cat("\n")
}))
```

# References
